---
title: "Green Bond Tracker — Data Cleaning: World Bank GSSS Bond Issuance"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This notebook documents the cleaning and processing of the World Bank / IFC
**Green, Social, Sustainability and Sustainability-Linked (GSSS) Bond Issuance**
dataset. Our goal is to extract **green bond issuance data only**, organize it
into a flat, map-ready table, and export a processed CSV that can be joined to
country boundary polygons in ArcGIS Pro.

## Data Source

- **Provider:** World Bank Data360 / International Finance Corporation (IFC)
- **Indicator:** `IFC_GB_GBAGG` — GSSS Bond Issuance
- **Original source:** LSEG (formerly Refinitiv); IMF staff calculations
- **Download URL:** <https://data360.worldbank.org/en/indicator/IFC_GB_GBAGG>
- **Date accessed:** `r Sys.Date()`
- **License:** World Bank Open Data (CC BY 4.0)

## Why Green Bonds Only?

The full dataset covers four ESG debt categories: Green, Social, Sustainability,
and Sustainability-Linked bonds. For this project, we focus exclusively on
**green bonds** — fixed-income instruments whose proceeds are directed to finance
or refinance projects with environmental benefits. This keeps our map focused
and our story clear: *Where in the world are green bonds being issued, and how
much capital are they mobilizing?*

# Setup

```{r load-libraries}
library(tidyverse)
library(knitr)
```

# 1. Load the Raw Data

The raw CSV lives in `data/raw/`. It is a wide, multi-dimensional dataset with
one row per combination of country × bond type × issuer type × currency ×
unit of measure × year.

```{r load-data}
raw <- read_csv("C:/Users/gabpe/OneDrive/Documents/src/green-bond-tracker-project/data/raw/IFC_GB_GBAGG.csv")

cat("Dimensions:", nrow(raw), "rows ×", ncol(raw), "columns\n")
```

Let's inspect the structure:

```{r inspect-structure}
glimpse(raw)
```

# 2. Understand the Dimensions

Before filtering, we need to understand the categorical breakdowns in this
dataset. Each "dimension" multiplies the row count.

```{r understand-dimensions}
cat("Bond Types:\n")
raw %>% distinct(BOND_TYPE, BOND_TYPE_LABEL) %>% kable()

cat("\nIssuer Types:\n")
raw %>% distinct(ISSUER_TYPE, ISSUER_TYPE_LABEL) %>% kable()

cat("\nCurrency Types:\n")
raw %>% distinct(CURRENCY, CURRENCY_LABEL) %>% kable()

cat("\nUnit of Measure:\n")
raw %>% distinct(UNIT_MEASURE, UNIT_MEASURE_LABEL, UNIT_MULT_LABEL) %>% kable()

cat("\nTime Period Range:", min(raw$TIME_PERIOD), "to", max(raw$TIME_PERIOD), "\n")
cat("Number of Entities (REF_AREA):", n_distinct(raw$REF_AREA), "\n")
```

# 3. Filter to Green Bonds Only

We apply the following filters:

1. **`BOND_TYPE == "GR"`** — Green bonds only.
2. **`ISSUER_TYPE == "_T"`** — All issuer types combined (we want the total per
   country, not the breakdown by who issued).
3. **`CURRENCY == "_T"`** — All currencies combined (we want the full picture
   regardless of whether bonds were issued in local or foreign currency).
4. **`UNIT_MEASURE == "USD"`** — Values in US dollars (billions). This is the
   only unit available for green bonds specifically; percentage-of-GDP data does
   not exist for the green-only subset.

```{r filter-green}
green_bonds <- raw %>%
  filter(
    BOND_TYPE == "GR",
    ISSUER_TYPE == "_T",
    CURRENCY == "_T",
    UNIT_MEASURE == "USD"
  )

cat("Rows after filtering:", nrow(green_bonds), "\n")
cat("Entities remaining:", n_distinct(green_bonds$REF_AREA), "\n")
cat("Year range:", min(green_bonds$TIME_PERIOD), "to", max(green_bonds$TIME_PERIOD), "\n")
```

# 4. Identify and Label Regional Aggregates

The dataset contains both individual countries (ISO 3166-1 alpha-3 codes) and
regional aggregates. We need to flag these so they can be handled appropriately.
Regional aggregates are useful context but **cannot be joined to country
polygons** on a map — they will live in a separate reference table.

```{r flag-regions}
# These codes represent regional aggregates, not individual countries
region_codes <- c("ECA", "LAC", "MNA", "SSA", "AMRC", "EAP_EX_CHN")

green_bonds <- green_bonds %>%
  mutate(is_region = REF_AREA %in% region_codes)

cat("Regional aggregates:", sum(green_bonds$is_region), "rows\n")
cat("Individual countries:", sum(!green_bonds$is_region), "rows\n")
```

# 5. Remove Zero-Issuance Entities

We only want entities (countries or regions) that have issued **at least one
green bond in at least one year**. If a country has zero green bond issuance
across all years, it is excluded entirely.

```{r remove-zeros}
# Identify entities with at least some green bond issuance
entities_with_issuance <- green_bonds %>%
  group_by(REF_AREA) %>%
  summarise(cumulative = sum(OBS_VALUE, na.rm = TRUE)) %>%
  filter(cumulative > 0) %>%
  pull(REF_AREA)

green_bonds <- green_bonds %>%
  filter(REF_AREA %in% entities_with_issuance)

cat("Entities with green bond issuance:", n_distinct(green_bonds$REF_AREA), "\n")
cat("  Countries:", green_bonds %>%
      filter(!is_region) %>%
      distinct(REF_AREA) %>%
      nrow(), "\n")
cat("  Regions:", green_bonds %>%
      filter(is_region) %>%
      distinct(REF_AREA) %>%
      nrow(), "\n")
```

# 6. Build the Processed Table

We now reshape the data into a flat, map-ready format. For each entity, we
compute:

- **Issuance for the most recent year (2024)** in USD billions
- **Cumulative all-time issuance (2012–2024)** in USD billions
- **First year of issuance** — when the country entered the green bond market
- **Number of active years** — how many years the country has issued green bonds

```{r build-table}
# Most recent year issuance
latest_year <- max(green_bonds$TIME_PERIOD)

issuance_latest <- green_bonds %>%
  filter(TIME_PERIOD == latest_year) %>%
  select(REF_AREA, REF_AREA_LABEL, OBS_VALUE, is_region) %>%
  rename(green_issuance_latest_usd_b = OBS_VALUE)

# Cumulative issuance
cumulative <- green_bonds %>%
  group_by(REF_AREA) %>%
  summarise(
    green_cumulative_usd_b = sum(OBS_VALUE, na.rm = TRUE),
    first_year_issued = min(TIME_PERIOD[OBS_VALUE > 0]),
    years_active = sum(OBS_VALUE > 0)
  )

# Join
processed <- issuance_latest %>%
  left_join(cumulative, by = "REF_AREA") %>%
  arrange(desc(green_cumulative_usd_b))

cat("Final processed table:", nrow(processed), "rows\n")
```

Let's preview the top entries:

```{r preview-top}
processed %>%
  select(
    ISO3 = REF_AREA,
    Country = REF_AREA_LABEL,
    Green_2024_USD_B = green_issuance_latest_usd_b,
    Green_Cumulative_USD_B = green_cumulative_usd_b,
    First_Year = first_year_issued,
    Years_Active = years_active,
    Is_Region = is_region
  ) %>%
  head(20) %>%
  kable(digits = 3, caption = "Top 20 Green Bond Issuers (Cumulative)")
```

# 7. Build the Time Series Table

For potential use in charts, dashboards, or ArcGIS time-enabled layers, we also
export a long-format time series of annual green bond issuance for all entities
that have ever issued.

```{r build-timeseries}
timeseries <- green_bonds %>%
  select(
    ISO3 = REF_AREA,
    Country = REF_AREA_LABEL,
    Year = TIME_PERIOD,
    Green_Issuance_USD_B = OBS_VALUE,
    Is_Region = is_region
  ) %>%
  arrange(ISO3, Year)

cat("Time series table:", nrow(timeseries), "rows\n")
cat("Covering", n_distinct(timeseries$ISO3), "entities across",
    n_distinct(timeseries$Year), "years\n")
```

# 8. Summary Statistics

Before exporting, let's document what we have.

```{r summary-stats}
cat("=== PROCESSED DATA SUMMARY ===\n\n")

cat("Total entities:", nrow(processed), "\n")
cat("  Countries:", sum(!processed$is_region), "\n")
cat("  Regional aggregates:", sum(processed$is_region), "\n\n")

cat("Latest year:", latest_year, "\n")
cat("Entities with issuance in", latest_year, ":",
    sum(processed$green_issuance_latest_usd_b > 0), "\n\n")

cat("Global cumulative green bond issuance (countries only):\n")
countries_only <- processed %>% filter(!is_region)
cat("  Total: $", round(sum(countries_only$green_cumulative_usd_b), 2), "B\n")
cat("  Mean per country: $",
    round(mean(countries_only$green_cumulative_usd_b), 3), "B\n")
cat("  Median per country: $",
    round(median(countries_only$green_cumulative_usd_b), 3), "B\n\n")

cat("Top 5 cumulative issuers:\n")
countries_only %>%
  slice_head(n = 5) %>%
  mutate(label = paste0("  ", REF_AREA_LABEL, ": $",
                        round(green_cumulative_usd_b, 2), "B")) %>%
  pull(label) %>%
  cat(sep = "\n")
```

# 9. Export

We export two CSV files:

1. **`green_bonds_by_country.csv`** — The flat, map-ready summary table (one row
   per entity). This is what gets joined to country polygons in ArcGIS Pro.
2. **`green_bonds_timeseries.csv`** — The annual time series for all entities.
   Useful for charts and time-enabled map layers.

```{r export}
# Rename columns for clarity in ArcGIS Pro (short, no spaces, descriptive)
export_summary <- processed %>%
  transmute(
    ISO3 = REF_AREA,
    Country = REF_AREA_LABEL,
    Green_2024_USD_B = round(green_issuance_latest_usd_b, 6),
    Green_Cumulative_USD_B = round(green_cumulative_usd_b, 6),
    First_Year = first_year_issued,
    Years_Active = years_active,
    Is_Region = is_region
  )


write_csv(export_summary, "../data/processed/green_bonds_by_country.csv")
write_csv(timeseries, "../data/processed/green_bonds_timeseries.csv")


cat("Exported:\n")
cat("  data/processed/green_bonds_by_country.csv —",
    nrow(export_summary), "rows\n")
cat("  data/processed/green_bonds_timeseries.csv —",
    nrow(timeseries), "rows\n")
```

# 10. Next Steps

With the cleaned data exported, the next phase of the project involves:

1. **Open ArcGIS Pro** and create a new Map project.
2. **Add a world countries boundary layer** from ArcGIS Living Atlas.
3. **Import `green_bonds_by_country.csv`** as a standalone table.
4. **Join the table to the country polygons** using the `ISO3` field.
5. **Symbolize the map** — graduated colors by `Green_Cumulative_USD_B` or
   `Green_2024_USD_B`.
6. **Add the IMF Carbon Footprint of Bank Loans layer** as a second thematic
   layer to provide the contrast analysis.
7. **Publish to ArcGIS Online** as a public web map.

---

*This notebook is part of the [Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).
Data sourced from the World Bank Data360 platform under CC BY 4.0 license.*
