---
title: "Green Bond Tracker — ISO Code Conversion for ArcGIS Pro"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

In the previous notebook (`01_clean_world_bank_gsss.Rmd`), we cleaned the raw
World Bank GSSS dataset down to green-bond-only issuance data. That produced two
files in `data/processed/`:

- `green_bonds_by_country.csv` — 75 rows (70 countries + 5 regional aggregates)
- `green_bonds_timeseries.csv` — annual time series for those 75 entities

When we loaded `green_bonds_by_country.csv` into ArcGIS Pro alongside the
**World Countries (Generalized)** boundary layer from Esri's Living Atlas, we
discovered a **code mismatch**: our data uses **ISO 3166-1 alpha-3** codes
(3 letters, e.g. `CHN`, `BRA`, `IND`) but the ArcGIS layer uses
**ISO 3166-1 alpha-2** codes (2 letters, e.g. `CN`, `BR`, `IN`).

This notebook adds an `ISO2` column to both processed files so they can be
joined to the ArcGIS country polygons. We also separate countries from regional
aggregates, since regional aggregates have no corresponding polygon to join to.

## Why Not Just Fix It Manually?

We could hand-edit the CSV or do a one-off conversion. But documenting the
transformation in code means:

1. **Reproducibility** — anyone can re-run this pipeline from raw data to
   map-ready output.
2. **Transparency** — the repo shows exactly how every field was derived.
3. **Auditability** — if a code mapping is wrong, we can trace it here.

# Setup

We use the `countrycode` package for the ISO3-to-ISO2 conversion. This package
maintains a curated lookup table of country code standards and handles edge cases
(name variations, historical changes, etc.) far more reliably than a hand-built
mapping table.

```{r load-libraries}
# install.packages("countrycode")  # Uncomment if not yet installed
# install.packages("tidyverse")    # Uncomment if not yet installed

library(tidyverse)
library(countrycode)
library(knitr)
```

# 1. Load the Processed Data

```{r load-data}
summary_df <- read_csv("../data/processed/green_bonds_by_country.csv")
timeseries_df <- read_csv("../data/processed/green_bonds_timeseries.csv")

cat("Summary table:", nrow(summary_df), "rows\n")
cat("Time series table:", nrow(timeseries_df), "rows\n")
```

Quick sanity check — confirm the data looks as expected:

```{r sanity-check}
summary_df %>%
  head(10) %>%
  kable(caption = "First 10 rows of summary table")
```

# 2. Separate Countries from Regional Aggregates

Regional aggregates (e.g., "Latin America & Caribbean") don't have ISO2 codes
and can't be joined to country polygons. We split them out now. They'll be
preserved in a separate reference file for use in charts or documentation, but
they won't go into the ArcGIS join.

```{r separate-regions}
regions_df <- summary_df %>% filter(Is_Region == TRUE)
countries_df <- summary_df %>% filter(Is_Region == FALSE)

cat("Countries:", nrow(countries_df), "\n")
cat("Regional aggregates:", nrow(regions_df), "\n\n")

cat("Regional aggregates being separated:\n")
regions_df %>%
  select(ISO3, Country) %>%
  kable()
```

# 3. Convert ISO3 to ISO2

The `countrycode()` function takes our 3-letter codes and returns the
corresponding 2-letter codes. We specify `origin = "iso3c"` (ISO alpha-3) and
`destination = "iso2c"` (ISO alpha-2).

```{r convert-codes}
countries_df <- countries_df %>%
  mutate(
    ISO2 = countrycode(
      sourcevar = ISO3,
      origin = "iso3c",
      destination = "iso2c"
    )
  )
```

## 3a. Verify the Conversion

This is the most important step. We check for any codes that failed to convert
(returned `NA`), and we eyeball a sample of mappings to make sure they're
correct.

```{r verify-conversion}
# Check for failed conversions
failed <- countries_df %>% filter(is.na(ISO2))

if (nrow(failed) > 0) {
  cat("WARNING:", nrow(failed), "codes failed to convert:\n")
  failed %>% select(ISO3, Country) %>% kable()
} else {
  cat("All", nrow(countries_df), "country codes converted successfully.\n")
}

# Check for duplicates (two ISO3 codes mapping to the same ISO2)
dupes <- countries_df %>%
  group_by(ISO2) %>%
  filter(n() > 1)

if (nrow(dupes) > 0) {
  cat("\nWARNING: Duplicate ISO2 codes detected:\n")
  dupes %>% select(ISO3, ISO2, Country) %>% kable()
} else {
  cat("No duplicate ISO2 codes. Each country maps uniquely.\n")
}
```

Let's also eyeball a sample of the mappings to make sure they pass the
common-sense test:

```{r eyeball-mappings}
countries_df %>%
  select(ISO3, ISO2, Country) %>%
  slice_sample(n = 15) %>%
  arrange(Country) %>%
  kable(caption = "Random sample of ISO3 → ISO2 mappings")
```

# 4. Reorganize Columns

We put `ISO2` right next to `ISO3` for clarity, and drop the `Is_Region` column
since this file now contains only countries.

```{r reorganize-columns}
countries_export <- countries_df %>%
  select(
    ISO3,
    ISO2,
    Country,
    Green_2024_USD_B,
    Green_Cumulative_USD_B,
    First_Year,
    Years_Active
  )

countries_export %>%
  head(15) %>%
  kable(digits = 3, caption = "Final country table (first 15 rows)")
```

# 5. Update the Time Series Too

The time series file also needs ISO2 codes for potential time-enabled layers in
ArcGIS. We apply the same conversion, and also split out regions.

```{r update-timeseries}
timeseries_countries <- timeseries_df %>%
  filter(Is_Region == FALSE) %>%
  mutate(
    ISO2 = countrycode(ISO3, "iso3c", "iso2c")
  ) %>%
  select(ISO3, ISO2, Country, Year, Green_Issuance_USD_B)

timeseries_regions <- timeseries_df %>%
  filter(Is_Region == TRUE)

cat("Time series — countries:", nrow(timeseries_countries), "rows\n")
cat("Time series — regions:", nrow(timeseries_regions), "rows\n")

# Verify no failed conversions
ts_failed <- timeseries_countries %>% filter(is.na(ISO2))
cat("Failed time series conversions:", nrow(ts_failed), "\n")
```

# 6. Summary Before Export

```{r final-summary}
cat("=== FINAL OUTPUT SUMMARY ===\n\n")

cat("Countries file:\n")
cat("  Rows:", nrow(countries_export), "\n")
cat("  Columns:", paste(names(countries_export), collapse = ", "), "\n")
cat("  Countries with 2024 issuance > 0:",
    sum(countries_export$Green_2024_USD_B > 0), "\n")
cat("  Total cumulative green issuance: $",
    round(sum(countries_export$Green_Cumulative_USD_B), 2), "B\n\n")

cat("Time series file:\n")
cat("  Rows:", nrow(timeseries_countries), "\n")
cat("  Entities:", n_distinct(timeseries_countries$ISO3), "\n")
cat("  Years:", min(timeseries_countries$Year), "to",
    max(timeseries_countries$Year), "\n\n")

cat("Regional aggregates file:\n")
cat("  Rows:", nrow(regions_df), "\n")
cat("  Entities:", n_distinct(regions_df$ISO3), "\n")
```

# 7. Export

We write three files to `data/processed/`:

1. **`green_bonds_by_country_iso2.csv`** — The map-ready file with ISO2 codes.
   This is what you join to country polygons in ArcGIS Pro.
2. **`green_bonds_timeseries_iso2.csv`** — The time series with ISO2 codes.
3. **`green_bonds_regional_aggregates.csv`** — Regional aggregates, kept
   separately for reference.

```{r export}
write_csv(countries_export, "../data/processed/green_bonds_by_country_iso2.csv")
write_csv(timeseries_countries, "../data/processed/green_bonds_timeseries_iso2.csv")
write_csv(regions_df, "../data/processed/green_bonds_regional_aggregates.csv")

cat("Exported:\n")
cat("  data/processed/green_bonds_by_country_iso2.csv —",
    nrow(countries_export), "rows\n")
cat("  data/processed/green_bonds_timeseries_iso2.csv —",
    nrow(timeseries_countries), "rows\n")
cat("  data/processed/green_bonds_regional_aggregates.csv —",
    nrow(regions_df), "rows\n")
```

# 8. Next Steps in ArcGIS Pro

With the new `green_bonds_by_country_iso2.csv` file, the ArcGIS Pro workflow is:

1. **Remove** the old `green_bonds_by_country.csv` standalone table from the
   project (right-click → Remove).
2. **Add Data** → navigate to `data/processed/green_bonds_by_country_iso2.csv`.
3. **Right-click** `World_Countries_Generalized` → **Joins and Relates** →
   **Add Join**.
4. Set **Input Join Field** to the 2-character ISO code column in the countries
   layer (the column titled `Two-character ISO Code for the Country`).
5. Set **Join Table** to `green_bonds_by_country_iso2.csv`.
6. Set **Join Table Field** to `ISO2`.
7. Click **OK** — the country polygons now carry your green bond data.

---

*This notebook is part of the
[Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).
Data sourced from the World Bank Data360 platform under CC BY 4.0 license.*
