---
title: "Green Bond Tracker — Data Cleaning: IMF Carbon Footprint of Bank Loans"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This notebook processes the IMF **Carbon Footprint of Bank Loans (CFBL)**
dataset for use as a second thematic layer in our ArcGIS Pro map. The CFBL
indicator quantifies how carbon-intensive a country's bank lending portfolio is.
By overlaying this with our green bond issuance data, we can ask a powerful
analytical question: *Are the countries with the dirtiest bank lending also the
ones issuing green bonds to finance the transition — or not?*

## Data Source

- **Provider:** International Monetary Fund (IMF), Climate Change Indicators
  Dashboard
- **Indicator:** Carbon Footprint-Adjusted Loans to Total Loans for Deposit
  Takers
- **Download page:** <https://climatedata.imf.org/datasets/596f11fea29d429ba6c5507e3756a751_0/explore>
- **ArcGIS Feature Service:** `https://services9.arcgis.com/weJ1QsnbMYJlCHdG/arcgis/rest/services/Indicator_9_Carbon_footprint_adjusted_loans_to_total_loans_for_deposit_takers/FeatureServer`
- **Original sources:** OECD Inter-Country Input-Output Database; IMF Statistics
  Department Questionnaire; IMF staff calculations
- **Date accessed:** `r Sys.Date()`

## What Does the Indicator Measure?

The CFBL takes the dollar value of bank loans to each economic sector, multiplies
each by that sector's carbon emissions factor (tonnes of CO2 per dollar of
output), sums across all sectors, and divides by total loans. A **higher ratio**
means a country's banking system is lending more heavily to carbon-intensive
industries. A **lower ratio** means lending is tilted toward cleaner sectors.

The dataset provides four variants of this indicator:

| Code   | Name                         | What It Captures                                      |
|--------|------------------------------|-------------------------------------------------------|
| ECFLI  | Emission Intensities         | Direct emissions only (raw ratio)                     |
| ECFLIN | Emission Intensities Norm.   | Same, but indexed to base year = 100                  |
| ECFLM  | Emission Multipliers         | Direct + indirect upstream emissions (raw ratio)      |
| ECFLMN | Emission Multipliers Norm.   | Same, but indexed to base year = 100                  |

**We will use `ECFLM` (Emission Multipliers, raw ratio)** for the map because:

1. It captures the **full carbon picture** including supply-chain emissions, not
   just direct combustion.
2. The **raw ratio** is directly comparable across countries (unlike the
   normalized version, which only shows change relative to each country's own
   base year).

# Setup

```{r load-libraries}
library(tidyverse)
library(knitr)
```

# 1. Load the Raw Data

```{r load-data}
raw <- read_csv("../data/raw/Carbon_Footprint_of_Bank_Loans.csv")

cat("Dimensions:", nrow(raw), "rows ×", ncol(raw), "columns\n")
cat("Countries:", n_distinct(raw$Country), "\n")
cat("Indicators:", n_distinct(raw$Indicator), "\n")
```

```{r inspect}
glimpse(raw)
```

# 2. Understand the Structure

The data is in **wide format**: each year is a separate column (`F2005` through
`F2018`). Each country appears four times — once per indicator variant. Let's
confirm this.

```{r structure-check}
raw %>%
  count(CTS_Code, Indicator) %>%
  kable(caption = "Rows per indicator variant")
```

And data availability grows over time (more countries report in later years):

```{r availability}
year_cols <- paste0("F", 2005:2018)

availability <- raw %>%
  filter(CTS_Code == "ECFLM") %>%
  summarise(across(all_of(year_cols), ~sum(!is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Year", values_to = "Countries") %>%
  mutate(Year = as.integer(str_remove(Year, "F")))

availability %>% kable(caption = "Countries with ECFLM data by year")
```

# 3. Filter to Emission Multipliers (ECFLM)

```{r filter-ecflm}
ecflm <- raw %>%
  filter(CTS_Code == "ECFLM")

cat("Rows after filtering to ECFLM:", nrow(ecflm), "\n")
cat("Countries:", n_distinct(ecflm$Country), "\n")
```

# 4. Reshape from Wide to Long

ArcGIS Pro works better with long-format data for time-enabled layers. We also
need long format to compute summary statistics per country.

```{r reshape}
ecflm_long <- ecflm %>%
  select(Country, ISO2, ISO3, all_of(year_cols)) %>%
  pivot_longer(
    cols = all_of(year_cols),
    names_to = "Year",
    values_to = "CFBL_Multiplier"
  ) %>%
  mutate(Year = as.integer(str_remove(Year, "F"))) %>%
  filter(!is.na(CFBL_Multiplier))

cat("Long-format rows:", nrow(ecflm_long), "\n")
cat("Countries:", n_distinct(ecflm_long$Country), "\n")
cat("Year range:", min(ecflm_long$Year), "to", max(ecflm_long$Year), "\n")
```

# 5. Build the Map-Ready Summary Table

For the static map layer, we need one row per country with the most recent
CFBL value (2018 for all 40 countries), plus contextual columns.

```{r build-summary}
cfbl_summary <- ecflm_long %>%
  group_by(Country, ISO2, ISO3) %>%
  summarise(
    CFBL_2018 = CFBL_Multiplier[Year == 2018],
    CFBL_First_Year = min(Year),
    CFBL_Years_Reported = n(),
    CFBL_Mean = mean(CFBL_Multiplier),
    CFBL_Trend = case_when(
      n() >= 3 ~ coef(lm(CFBL_Multiplier ~ Year))[2],
      TRUE ~ NA_real_
    ),
    .groups = "drop"
  ) %>%
  arrange(desc(CFBL_2018))

cat("Summary table:", nrow(cfbl_summary), "rows\n")
```

The `CFBL_Trend` column is the linear slope (change in CFBL per year). A
**negative trend** means the country's bank lending is getting cleaner over time.
A **positive trend** means it's getting dirtier.

```{r preview-summary}
cfbl_summary %>%
  head(20) %>%
  kable(digits = 2, caption = "Top 20 countries by carbon footprint of bank loans (2018)")
```

# 6. Identify Overlap with Green Bond Data

This is the analytical heart of the project. Which countries appear in both
datasets?

```{r overlap-analysis}
green <- read_csv("../data/processed/green_bonds_by_country.csv") %>%
  filter(Is_Region == FALSE)

overlap <- inner_join(
  cfbl_summary %>% select(ISO3, Country, CFBL_2018, CFBL_Trend),
  green %>% select(ISO3, Green_Cumulative_USD_B, Green_2024_USD_B),
  by = "ISO3"
)

cat("Countries with BOTH green bond and carbon footprint data:", nrow(overlap), "\n\n")

overlap %>%
  arrange(desc(CFBL_2018)) %>%
  kable(digits = 2,
        caption = "Countries with both CFBL and green bond data — sorted by carbon intensity")
```

This table is where the interesting stories emerge. For instance:

- **Kazakhstan** has the highest CFBL (1205) but minimal green bond issuance
  ($0.12B) — a potential gap in transition finance.
- **Indonesia** has a high CFBL (645) and meaningful green bond issuance
  ($11.4B) — possibly financing the transition.
- **Czechia** and **Hungary** have moderate CFBL scores but strong green bond
  programs — potentially proactive.

# 7. Export

We export three files:

1. **`cfbl_by_country.csv`** — One row per country, map-ready for ArcGIS Pro
   join.
2. **`cfbl_timeseries.csv`** — Long-format annual data for time-enabled layers
   or charts.
3. **`cfbl_green_bond_overlap.csv`** — The combined dataset for countries in
   both sources.

```{r export}
write_csv(cfbl_summary, "../data/processed/cfbl_by_country.csv")
write_csv(ecflm_long, "../data/processed/cfbl_timeseries.csv")
write_csv(overlap, "../data/processed/cfbl_green_bond_overlap.csv")

cat("Exported:\n")
cat("  ../data/processed/cfbl_by_country.csv —",
    nrow(cfbl_summary), "rows\n")
cat("  ../data/processed/cfbl_timeseries.csv —",
    nrow(ecflm_long), "rows\n")
cat("  ../data/processed/cfbl_green_bond_overlap.csv —",
    nrow(overlap), "rows\n")
```

# 8. Next Steps in ArcGIS Pro

With `cfbl_by_country.csv` ready:

1. **Add Data** → `data/processed/cfbl_by_country.csv` as a standalone table.
2. **Right-click** `World_Countries_Generalized` → **Joins and Relates** →
   **Add Join**.
3. Join on `ISO2` (the 2-character ISO code column in the countries layer) to
   `ISO2` in the CFBL table.

**Important:** ArcGIS Pro only supports one active join per field at a time. Since
we already joined the green bond data on the same ISO2 field, we have two
options:

- **Option A:** Export the countries layer to a new feature class (right-click →
  Data → Export Features) to make the green bond join permanent, then add the
  CFBL join on top.
- **Option B:** Merge both CSVs into a single table before joining, so only one
  join is needed.

Option B is cleaner. We will build that combined table in the next notebook if
needed.

---

*This notebook is part of the
[Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).
IMF data accessed from the Climate Change Indicators Dashboard.*
