---
title: "Green Bond Tracker — Data Cleaning: IMF ESG Bond Issuances"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This notebook processes the IMF Climate Data Dashboard's **ESG Bond Issuances**
dataset — a comprehensive source originally from LSEG (formerly Refinitiv). This
dataset covers **109 countries** including many advanced economies that our World
Bank dataset (focused on emerging markets) does not. It also provides sovereign
green bond issuances as a separate indicator, which tells us specifically about
government-led green finance commitments.

## Why This Dataset Matters

Our World Bank GSSS dataset (notebook 01) covers **70 countries**, mostly
emerging markets and developing economies. This IMF dataset adds **41 additional
countries** — including the United States, Germany, France, the UK, Japan,
Australia, and other major issuers. Combining the two gives us the most complete
global picture available.

## Data Source

- **Provider:** IMF Climate Change Indicators Dashboard / LSEG (Refinitiv)
- **Download page:** <https://climatedata.imf.org/datasets/8e2772e0b65f4e33a80183ce9583d062_0/explore>
- **Date accessed:** `r Sys.Date()`
- **Time coverage:** 2006–2024
- **Units:** Billion US Dollars

# Setup

```{r load-libraries}
# install.packages("countrycode")  # Uncomment if not yet installed
library(tidyverse)
library(countrycode)
library(knitr)
```

# 1. Load the Raw Data

```{r load-data}
raw <- read_csv("../data/raw/Green_Bonds.csv")

cat("Dimensions:", nrow(raw), "rows ×", ncol(raw), "columns\n")
cat("Countries:", n_distinct(raw$Country), "\n")
cat("Indicators:", n_distinct(raw$Indicator), "\n")
```

# 2. Understand the Indicators

This dataset is much richer than the World Bank one — it contains 25 different
indicators covering issuances, outstanding amounts, sovereign breakdowns, and
currency splits.

```{r list-indicators}
raw %>%
  count(Indicator) %>%
  arrange(desc(n)) %>%
  kable(caption = "Available indicators and row counts")
```

For our green bond map, we care about these indicators:

1. **`Green Bond Issuances by Country`** — Total green bond issuance per country
   per year. This is our primary layer.
2. **`Sovereign Green Bond Issuances`** — Green bonds issued specifically by
   governments. This enriches the story.

The other indicators (social bonds, sustainability bonds, outstanding amounts,
currency splits, issuer-type breakdowns) are available for future analysis but
fall outside our green-bond-only scope.

# 3. Extract Green Bond Issuances by Country

```{r filter-green-by-country}
year_cols <- paste0("F", 2006:2024)

green_by_country <- raw %>%
  filter(Indicator == "Green Bond Issuances by Country") %>%
  select(Country, ISO2, ISO3, all_of(year_cols))

cat("Countries with green bond issuance data:", nrow(green_by_country), "\n")
```

## 3a. Reshape to Long Format

```{r reshape-green}
green_long <- green_by_country %>%
  pivot_longer(
    cols = all_of(year_cols),
    names_to = "Year",
    values_to = "Green_Issuance_USD_B"
  ) %>%
  mutate(Year = as.integer(str_remove(Year, "F"))) %>%
  replace_na(list(Green_Issuance_USD_B = 0))

cat("Long-format rows:", nrow(green_long), "\n")
```

## 3b. Build Country Summary

```{r green-summary}
imf_green_summary <- green_long %>%
  group_by(Country, ISO2, ISO3) %>%
  summarise(
    IMF_Green_2024_USD_B = sum(Green_Issuance_USD_B[Year == 2024]),
    IMF_Green_Cumulative_USD_B = sum(Green_Issuance_USD_B),
    IMF_Green_First_Year = ifelse(
      any(Green_Issuance_USD_B > 0),
      min(Year[Green_Issuance_USD_B > 0]),
      NA_integer_
    ),
    IMF_Green_Years_Active = sum(Green_Issuance_USD_B > 0),
    .groups = "drop"
  ) %>%
  filter(IMF_Green_Cumulative_USD_B > 0) %>%
  arrange(desc(IMF_Green_Cumulative_USD_B))

cat("Countries with any green bond issuance:", nrow(imf_green_summary), "\n")
```

```{r preview-green}
imf_green_summary %>%
  head(20) %>%
  kable(digits = 3, caption = "Top 20 green bond issuers (IMF/LSEG data)")
```

# 4. Extract Sovereign Green Bond Issuances

Sovereign green bonds are a subset — bonds issued by central governments. This
tells us which governments are putting their money where their mouth is on
climate policy.

```{r filter-sovereign}
sovereign <- raw %>%
  filter(Indicator == "Sovereign Green Bond Issuances") %>%
  select(Country, ISO2, ISO3, all_of(year_cols))

sovereign_long <- sovereign %>%
  pivot_longer(
    cols = all_of(year_cols),
    names_to = "Year",
    values_to = "Sovereign_Green_USD_B"
  ) %>%
  mutate(Year = as.integer(str_remove(Year, "F"))) %>%
  replace_na(list(Sovereign_Green_USD_B = 0))

sovereign_summary <- sovereign_long %>%
  group_by(Country, ISO2, ISO3) %>%
  summarise(
    Sovereign_Green_Cumulative_USD_B = sum(Sovereign_Green_USD_B),
    Sovereign_First_Year = ifelse(
      any(Sovereign_Green_USD_B > 0),
      min(Year[Sovereign_Green_USD_B > 0]),
      NA_integer_
    ),
    .groups = "drop"
  ) %>%
  filter(Sovereign_Green_Cumulative_USD_B > 0) %>%
  arrange(desc(Sovereign_Green_Cumulative_USD_B))

cat("Countries with sovereign green bonds:", nrow(sovereign_summary), "\n")
```

```{r preview-sovereign}
sovereign_summary %>%
  head(15) %>%
  kable(digits = 3, caption = "Top 15 sovereign green bond issuers")
```

# 5. Merge Green and Sovereign Data

We combine both into a single country-level table so each country shows total
green issuance alongside sovereign green issuance.

```{r merge-green-sovereign}
imf_combined <- imf_green_summary %>%
  left_join(
    sovereign_summary %>%
      select(ISO3, Sovereign_Green_Cumulative_USD_B, Sovereign_First_Year),
    by = "ISO3"
  ) %>%
  replace_na(list(Sovereign_Green_Cumulative_USD_B = 0)) %>%
  mutate(
    Sovereign_Share_Pct = round(
      Sovereign_Green_Cumulative_USD_B / IMF_Green_Cumulative_USD_B * 100, 1
    )
  )

cat("Final combined table:", nrow(imf_combined), "countries\n")
```

```{r preview-combined}
imf_combined %>%
  head(20) %>%
  kable(digits = 3,
        caption = "Top 20 — total green issuance with sovereign share")
```

# 6. Compare with World Bank Data

```{r compare-wb}
wb <- read_csv("../data/processed/green_bonds_by_country.csv") %>%
  filter(Is_Region == FALSE)

imf_iso3 <- imf_combined$ISO3
wb_iso3 <- wb$ISO3

cat("=== COVERAGE COMPARISON ===\n")
cat("IMF/LSEG countries:", length(imf_iso3), "\n")
cat("World Bank/IFC countries:", length(wb_iso3), "\n")
cat("In both:", length(intersect(imf_iso3, wb_iso3)), "\n")
cat("Only in IMF:", length(setdiff(imf_iso3, wb_iso3)), "\n")
cat("Only in World Bank:", length(setdiff(wb_iso3, imf_iso3)), "\n")
cat("Combined unique countries:", length(union(imf_iso3, wb_iso3)), "\n")
```

The IMF dataset adds major advanced economies. Together the two datasets give us
the most complete global picture of green bond issuance.

# 7. Export

```{r export}
write_csv(imf_combined, "../data/processed/imf_green_bonds_by_country.csv")
write_csv(green_long, "../data/processed/imf_green_bonds_timeseries.csv")

cat("Exported:\n")
cat("  ../data/processed/imf_green_bonds_by_country.csv —",
    nrow(imf_combined), "rows\n")
cat("  ../data/processed/imf_green_bonds_timeseries.csv —",
    nrow(green_long), "rows\n")
```

# 8. Next Steps

The IMF/LSEG and World Bank/IFC datasets overlap for 49 countries but each has
unique entries. In the final combination notebook (`06_combine_for_arcgis.Rmd`),
we will merge all sources into one master table for the ArcGIS Pro join.

---

*This notebook is part of the
[Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).
Data sourced from the IMF Climate Change Indicators Dashboard (LSEG/Refinitiv).*
