---
title: "Green Bond Tracker — Data Cleaning: Climate Bonds Initiative Certified Bonds"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This notebook processes the **Climate Bonds Initiative (CBI) Certified Climate
Bonds Database** — a bond-level dataset of instruments that have been
independently verified and certified under the Climate Bonds Standard. Unlike our
other datasets which are country-level aggregates, this one lists **individual
bonds and loans** with details on the issuer, sector, size, and verification
status.

## Why This Dataset Matters

The CBI database tells a different part of the story. The World Bank and IMF
datasets answer "how much green debt is a country issuing?" This dataset answers
"what specifically is that money being used for?" — low-carbon transport, solar,
wind, buildings, water infrastructure, etc. It also adds a quality dimension:
these bonds have been **independently verified**, which is a higher bar than
self-labeling.

## Data Source

- **Provider:** Climate Bonds Initiative
- **Download page:** <https://www.climatebonds.net/data-insights/market-data/certified-climate-bonds-database>
- **Date accessed:** `r Sys.Date()`

# Setup

```{r load-libraries}
# install.packages("countrycode")  # Uncomment if not yet installed
library(tidyverse)
library(countrycode)
library(knitr)
```

# 1. Load the Raw Data

```{r load-data}
raw <- read_csv("../data/raw/bonds_export.csv")

cat("Dimensions:", nrow(raw), "rows ×", ncol(raw), "columns\n")
cat("Unique issuers:", n_distinct(raw$`Issuer / Applicant`), "\n")
cat("Unique countries:", n_distinct(raw$`Issuer Country`, na.rm = TRUE), "\n")
cat("Total bonds/loans:", nrow(raw), "\n")
```

```{r inspect}
glimpse(raw)
```

# 2. Data Quality Issues

This dataset has several issues we need to address before aggregation.

## 2a. Missing and Malformed Country Names

```{r country-issues}
# Check for null countries
null_countries <- sum(is.na(raw$`Issuer Country`))
cat("Rows with missing country:", null_countries, "\n\n")

# List all unique country names
cat("All unique country names:\n")
sort(unique(raw$`Issuer Country`[!is.na(raw$`Issuer Country`)])) %>%
  paste("  ", .) %>%
  cat(sep = "\n")
```

Issues identified:

- **`"UAE"` and `"United Arab Emirates"`** refer to the same country.
- **`"Hong KongGermany"`** and **`"Hong KongChina"`** are malformed entries
  (likely data entry errors where a multi-listing separator was missing).
- **`"Ivory Coast"`** is the common name; ISO standard is Côte d'Ivoire.
- **`"South Korea"`** should be mapped to Korea, Rep. of.
- **`"Europe"`** and **`"No single country  covers the European Economic Area"`**
  are not individual countries.
- **4 rows** have `NA` for country.

## 2b. Fix Country Names

```{r fix-countries}
cbi <- raw %>%
  mutate(
    `Issuer Country` = case_when(
      `Issuer Country` == "UAE" ~ "United Arab Emirates",
      `Issuer Country` == "Hong KongGermany" ~ "Hong Kong",
      `Issuer Country` == "Hong KongChina" ~ "Hong Kong",
      `Issuer Country` == "Ivory Coast" ~ "Cote d'Ivoire",
      `Issuer Country` == "South Korea" ~ "Korea, Republic of",
      TRUE ~ `Issuer Country`
    )
  ) %>%
  # Remove rows that aren't assignable to a single country
  filter(
    !is.na(`Issuer Country`),
    !`Issuer Country` %in% c(
      "Europe",
      "No single country  covers the European Economic Area"
    )
  )

cat("Rows after cleaning:", nrow(cbi), "\n")
cat("Unique countries after cleaning:", n_distinct(cbi$`Issuer Country`), "\n")
```

## 2c. Add ISO Codes

The CBI dataset uses full country names, not ISO codes. We use the `countrycode`
package to map them.

```{r add-iso}
cbi <- cbi %>%
  mutate(
    ISO2 = countrycode(
      `Issuer Country`, origin = "country.name", destination = "iso2c"
    ),
    ISO3 = countrycode(
      `Issuer Country`, origin = "country.name", destination = "iso3c"
    )
  )

# Check for failed conversions
failed <- cbi %>% filter(is.na(ISO2))
if (nrow(failed) > 0) {
  cat("WARNING: Failed to convert", n_distinct(failed$`Issuer Country`),
      "country names:\n")
  failed %>% distinct(`Issuer Country`) %>% pull() %>% cat(sep = "\n")
} else {
  cat("All country names converted successfully.\n")
}
```

## 2d. Handle Bond Status

Some bonds have been revoked, redeemed, or discontinued. We flag these but keep
them in the dataset for transparency.

```{r status-check}
cbi %>%
  count(Status) %>%
  kable(caption = "Bond status distribution")
```

# 3. Clean the Size Column

The USD equivalent size column is already numeric but has some zero values
(likely pending or not yet disclosed).

```{r clean-size}
cbi <- cbi %>%
  mutate(Size_USD_B = `Size (USD equivalent)` / 1e9)

cat("Bonds with non-zero size:", sum(cbi$Size_USD_B > 0, na.rm = TRUE), "\n")
cat("Bonds with zero or NA size:", sum(cbi$Size_USD_B == 0 | is.na(cbi$Size_USD_B)), "\n")
cat("Total value (non-zero):", round(sum(cbi$Size_USD_B[cbi$Size_USD_B > 0],
                                         na.rm = TRUE), 2), "B USD\n")
```

# 4. Standardize Sector Names

The sector criteria column has inconsistencies (e.g., extra spaces, mixed
delimiters for multi-sector bonds).

```{r clean-sectors}
cbi <- cbi %>%
  mutate(
    Sector_Clean = `Sector Criteria` %>%
      str_replace_all("\\s{2,}", " ") %>%
      str_trim()
  )

# For multi-sector bonds, extract the primary (first listed) sector
cbi <- cbi %>%
  mutate(
    Primary_Sector = str_extract(Sector_Clean, "^[A-Za-z ]+") %>%
      str_trim() %>%
      str_to_title()
  )

cbi %>%
  count(Primary_Sector) %>%
  arrange(desc(n)) %>%
  head(12) %>%
  kable(caption = "Bonds by primary sector")
```

# 5. Aggregate by Country

For the map layer, we need one row per country with summary statistics.

```{r aggregate-country}
cbi_by_country <- cbi %>%
  group_by(`Issuer Country`, ISO2, ISO3) %>%
  summarise(
    CBI_Bond_Count = n(),
    CBI_Total_USD_B = sum(Size_USD_B, na.rm = TRUE),
    CBI_Unique_Issuers = n_distinct(`Issuer / Applicant`),
    CBI_First_Issue = min(`Issue date`, na.rm = TRUE),
    CBI_Latest_Issue = max(`Issue date`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(CBI_Total_USD_B))

cat("Countries:", nrow(cbi_by_country), "\n")
```

```{r preview-country}
cbi_by_country %>%
  head(20) %>%
  kable(digits = 3, caption = "Top 20 countries by CBI-certified green bond value")
```

# 6. Aggregate by Sector (Global)

This is useful reference data — what sectors is certified green finance going to?

```{r aggregate-sector}
cbi_by_sector <- cbi %>%
  group_by(Primary_Sector) %>%
  summarise(
    Bond_Count = n(),
    Total_USD_B = sum(Size_USD_B, na.rm = TRUE),
    Countries = n_distinct(ISO3),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_USD_B))

cbi_by_sector %>%
  kable(digits = 3, caption = "CBI-certified bonds by sector (global)")
```

# 7. Aggregate by Country × Sector

For a richer map experience (pop-ups showing sector breakdown per country):

```{r country-sector}
cbi_country_sector <- cbi %>%
  group_by(ISO2, ISO3, `Issuer Country`, Primary_Sector) %>%
  summarise(
    Bond_Count = n(),
    Total_USD_B = sum(Size_USD_B, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(ISO3, desc(Total_USD_B))

cat("Country × Sector combinations:", nrow(cbi_country_sector), "\n")
```

# 8. Export

```{r export}
write_csv(cbi_by_country, "../data/processed/cbi_by_country.csv")
write_csv(cbi_by_sector, "../data/processed/cbi_by_sector.csv")
write_csv(cbi_country_sector, "../data/processed/cbi_country_sector.csv")

cat("Exported:\n")
cat("  ../data/processed/cbi_by_country.csv —",
    nrow(cbi_by_country), "rows\n")
cat("  ../data/processed/cbi_by_sector.csv —",
    nrow(cbi_by_sector), "rows\n")
cat("  ../data/processed/cbi_country_sector.csv —",
    nrow(cbi_country_sector), "rows\n")
```

# 9. Key Observations

The CBI dataset adds a **quality dimension** to our map. While the World Bank
and IMF datasets measure the total volume of self-labeled green bonds, the CBI
database only includes bonds that have been independently verified under the
Climate Bonds Standard. This means:

- A country may have large green bond issuance (IMF/World Bank) but few
  CBI-certified bonds — suggesting much of their green labeling is self-declared.
- A country with a high ratio of CBI-certified to total green bonds has a more
  credible green finance market.
- The **sector breakdown** reveals what the money is actually financing, which
  matters enormously for climate impact assessment.

---

*This notebook is part of the
[Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).
Data sourced from the Climate Bonds Initiative Certified Climate Bonds Database.*
