---
title: "Green Bond Tracker — Combine All Sources for ArcGIS Pro"
author: "Green Bond Tracker Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This is the final data preparation notebook. It merges all four processed
datasets into a **single master CSV** that can be joined to country polygons in
ArcGIS Pro with one join operation. No more stacking joins — one file, one join,
all data.

## Datasets Being Combined

| # | Source | Notebook | Countries | Key Fields |
|---|--------|----------|-----------|------------|
| 1 | World Bank / IFC | 01, 02 | 70 | Green bond issuance (emerging markets focus) |
| 2 | IMF / LSEG | 04 | 90 | Green bond issuance (global, incl. advanced economies) + sovereign |
| 3 | IMF | 03 | 40 | Carbon footprint of bank loans |
| 4 | Climate Bonds Initiative | 05 | 47 | Certified bonds (count, value, sectors) |

Combined, these cover **114 unique countries**.

## Handling Overlap

The World Bank and IMF green bond datasets overlap for ~49 countries. They come
from different providers (IFC vs. LSEG/Refinitiv) and use slightly different
methodologies, so values will not match exactly. Our approach:

- We keep **both** sets of values in separate columns so users can compare.
- We create a **Best Estimate** column that takes the IMF value when available
  (broader coverage, includes advanced economies), falling back to the World Bank
  value for countries only in that dataset.

# Setup

```{r load-libraries}
library(tidyverse)
library(knitr)
```

# 1. Load All Processed Datasets

**Important:** Namibia's ISO2 code is literally `"NA"`, which R's `read_csv`
interprets as a missing value. We prevent this with `na = c("")` so that only
empty strings are treated as NA.

```{r load-data}
wb <- read_csv(
  "../data/processed/green_bonds_by_country_iso2.csv",
  na = c("")
)
imf <- read_csv(
  "../data/processed/imf_green_bonds_by_country.csv",
  na = c("")
)
cfbl <- read_csv(
  "../data/processed/cfbl_by_country.csv",
  na = c("")
)
cbi <- read_csv(
  "../data/processed/cbi_by_country.csv",
  na = c("")
)

cat("World Bank:", nrow(wb), "countries\n")
cat("IMF Green Bonds:", nrow(imf), "countries\n")
cat("CFBL:", nrow(cfbl), "countries\n")
cat("CBI:", nrow(cbi), "countries\n")
```

Quick Namibia check:

```{r namibia-check}
cat("Namibia in WB — ISO2:", wb$ISO2[wb$ISO3 == "NAM"], "\n")
cat("Namibia in IMF — ISO2:", imf$ISO2[imf$ISO3 == "NAM"], "\n")
```

If Namibia's ISO2 still shows as `NA` (missing), we fix it manually:

```{r fix-namibia}
wb$ISO2[wb$ISO3 == "NAM"] <- "NA"
imf$ISO2[imf$ISO3 == "NAM"] <- "NA"
```

# 2. Build the Master Country List

We start from the union of all ISO2 codes across all four datasets, then join
each source onto that scaffold.

```{r build-scaffold}
# Collect all ISO2-ISO3-Country combinations from all sources
all_countries <- bind_rows(
  wb %>% select(ISO2, ISO3, Country),
  imf %>% select(ISO2, ISO3, Country),
  cfbl %>% select(ISO2, ISO3, Country),
  cbi %>% select(ISO2, ISO3, Country = `Issuer Country`)
) %>%
  # Prefer IMF country names (most standardized), then WB, then CBI
  group_by(ISO2) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  arrange(ISO2)

cat("Master country list:", nrow(all_countries), "unique countries\n")
```

# 3. Join All Sources

```{r join-all}
master <- all_countries %>%
  # World Bank green bond data
  left_join(
    wb %>% select(
      ISO2,
      WB_Green_2024 = Green_2024_USD_B,
      WB_Green_Cumul = Green_Cumulative_USD_B,
      WB_First_Year = First_Year,
      WB_Years_Active = Years_Active
    ),
    by = "ISO2"
  ) %>%
  # IMF/LSEG green bond data
  left_join(
    imf %>% select(
      ISO2,
      IMF_Green_2024 = IMF_Green_2024_USD_B,
      IMF_Green_Cumul = IMF_Green_Cumulative_USD_B,
      Sov_Green_Cumul = Sovereign_Green_Cumulative_USD_B,
      Sov_Share_Pct = Sovereign_Share_Pct
    ),
    by = "ISO2"
  ) %>%
  # Carbon footprint of bank loans
  left_join(
    cfbl %>% select(
      ISO2,
      CFBL_2018,
      CFBL_Trend,
      CFBL_Mean
    ),
    by = "ISO2"
  ) %>%
  # CBI certified bonds
  left_join(
    cbi %>% select(
      ISO2,
      CBI_Count = CBI_Bond_Count,
      CBI_USD_B = CBI_Total_USD_B,
      CBI_Issuers = CBI_Unique_Issuers
    ),
    by = "ISO2"
  )

cat("Master table:", nrow(master), "rows ×", ncol(master), "columns\n")
```

# 4. Create Best-Estimate Columns

For the map's primary symbology, we need a single "green bond issuance" value
per country. We prefer the IMF data (broader coverage) and fall back to World
Bank where IMF is missing.

```{r best-estimate}
master <- master %>%
  mutate(
    Green_2024_Best = coalesce(IMF_Green_2024, WB_Green_2024),
    Green_Cumul_Best = coalesce(IMF_Green_Cumul, WB_Green_Cumul),
    Data_Sources = paste0(
      ifelse(!is.na(WB_Green_Cumul), "WB", ""),
      ifelse(!is.na(IMF_Green_Cumul), "+IMF", ""),
      ifelse(!is.na(CFBL_2018), "+CFBL", ""),
      ifelse(!is.na(CBI_Count), "+CBI", "")
    ) %>% str_remove("^\\+")
  )
```

# 5. Add Coverage Flags

Boolean flags make it easy to filter in ArcGIS Pro.

```{r flags}
master <- master %>%
  mutate(
    Has_WB = !is.na(WB_Green_Cumul),
    Has_IMF = !is.na(IMF_Green_Cumul),
    Has_CFBL = !is.na(CFBL_2018),
    Has_CBI = !is.na(CBI_Count),
    Source_Count = Has_WB + Has_IMF + Has_CFBL + Has_CBI
  )

cat("\n=== DATA COVERAGE ===\n")
cat("Countries with World Bank data:", sum(master$Has_WB), "\n")
cat("Countries with IMF data:", sum(master$Has_IMF), "\n")
cat("Countries with CFBL data:", sum(master$Has_CFBL), "\n")
cat("Countries with CBI data:", sum(master$Has_CBI), "\n")
cat("Countries with 1 source:", sum(master$Source_Count == 1), "\n")
cat("Countries with 2 sources:", sum(master$Source_Count == 2), "\n")
cat("Countries with 3 sources:", sum(master$Source_Count == 3), "\n")
cat("Countries with all 4 sources:", sum(master$Source_Count == 4), "\n")
```

# 6. Final Column Order and Preview

```{r reorder}
master_export <- master %>%
  select(
    # Identity
    ISO2, ISO3, Country,
    # Best estimate (use for primary symbology)
    Green_2024_Best, Green_Cumul_Best,
    # World Bank detail
    WB_Green_2024, WB_Green_Cumul, WB_First_Year, WB_Years_Active,
    # IMF/LSEG detail
    IMF_Green_2024, IMF_Green_Cumul, Sov_Green_Cumul, Sov_Share_Pct,
    # Carbon footprint
    CFBL_2018, CFBL_Trend, CFBL_Mean,
    # CBI certified
    CBI_Count, CBI_USD_B, CBI_Issuers,
    # Metadata
    Data_Sources, Source_Count
  ) %>%
  arrange(desc(Green_Cumul_Best))

cat("Final master table:", nrow(master_export), "rows ×",
    ncol(master_export), "columns\n")
```

```{r preview-top}
master_export %>%
  head(25) %>%
  kable(digits = 3, caption = "Top 25 countries by best-estimate cumulative green issuance")
```

# 7. Summary Statistics

```{r summary}
cat("=== FINAL SUMMARY ===\n\n")

cat("Total countries:", nrow(master_export), "\n\n")

cat("Green bond issuance (best estimate):\n")
cat("  Global cumulative: $",
    round(sum(master_export$Green_Cumul_Best, na.rm = TRUE), 1), "B\n")
cat("  Global 2024: $",
    round(sum(master_export$Green_2024_Best, na.rm = TRUE), 1), "B\n")
cat("  Countries with any issuance:",
    sum(master_export$Green_Cumul_Best > 0, na.rm = TRUE), "\n\n")

cat("Sovereign green bonds:\n")
cat("  Countries with sovereign issuance:",
    sum(master_export$Sov_Green_Cumul > 0, na.rm = TRUE), "\n")
cat("  Total sovereign: $",
    round(sum(master_export$Sov_Green_Cumul, na.rm = TRUE), 1), "B\n\n")

cat("Carbon footprint of bank loans:\n")
cat("  Countries with CFBL data:", sum(!is.na(master_export$CFBL_2018)), "\n")
cat("  Highest: ",
    master_export$Country[which.max(master_export$CFBL_2018)], "(",
    round(max(master_export$CFBL_2018, na.rm = TRUE), 0), ")\n")
cat("  Lowest: ",
    master_export$Country[which.min(master_export$CFBL_2018)], "(",
    round(min(master_export$CFBL_2018, na.rm = TRUE), 0), ")\n\n")

cat("CBI certified bonds:\n")
cat("  Countries with certified bonds:", sum(!is.na(master_export$CBI_Count)), "\n")
cat("  Total certified bonds:", sum(master_export$CBI_Count, na.rm = TRUE), "\n")
cat("  Total certified value: $",
    round(sum(master_export$CBI_USD_B, na.rm = TRUE), 1), "B\n")
```

# 8. Export

```{r export}
write_csv(master_export, "../data/processed/map_data_combined.csv",
          na = "")

cat("Exported:\n")
cat("  ../data/processed/map_data_combined.csv —",
    nrow(master_export), "rows ×", ncol(master_export), "columns\n")
```

# 9. ArcGIS Pro Instructions

With `map_data_combined.csv` in hand, here's the complete ArcGIS workflow:

1. **Remove all existing joins** from `World_Countries_Generalized`:
   - Right-click the layer → **Joins and Relates** → **Remove All Joins**
2. **Remove the old standalone table** (`green_bonds_by_country_iso2.csv`)
   from the project.
3. **Add Data** → navigate to
   `data/processed/map_data_combined.csv`.
4. **Right-click** `World_Countries_Generalized` → **Joins and Relates** →
   **Add Join**:
   - **Input Join Field:** `Two-character ISO Code for the Country`
   - **Join Table:** `map_data_combined.csv`
   - **Join Table Field:** `ISO2`
5. **Symbolize** with Graduated Colors on `Green_Cumul_Best` for the
   primary green bond story.
6. Optionally add a second layer or use the `CFBL_2018` field for a bivariate
   map comparing green bond issuance to carbon footprint.

---

*This notebook is part of the
[Green Bond Tracker Project](https://github.com/YOUR_USERNAME/green-bond-tracker-project).*
